---
title: "Analysis of growth curve outputs"
author: "Saaz Sakrikar"
date: "2025-1-4"
output: html_document
---

This code takes in manually combined growthcurver output from all previous growth measurements. I am using only haploid control strains DGY1, DGY2807, DGY500, DGY1315, and DGY3183 here. 

It evaluates whether growth rate or area under curve is a more consistent and meaningful metric. It evaluates looks for outlier batches, where all replicates from a given batch are out-of-whack with all other batches.

Having found the correct metric (AUC, because of 34C and 37C high FLC data which becomes nonsensical when looked at by growth rate), and removed outliers, it plots a fitness landscape for S cerevisiae across 3 temps (30, 34, 37C) and 7 fluconazole concs (0, 4, 8, 16, 32, 64, 128mg/uL). This is part of Fig. 1.

```{r setup, warning=FALSE}
library(ggplot2)
library(readxl)
library(forcats)
library(tidyverse)

All_growthcurve_output = read_excel("Control_growthcurve_output_2026Jan9.xlsx")
```


```{r overview_plot}
#separate by type - haploid and diploid - and temperature, within each type
Haploids=All_growthcurve_output[All_growthcurve_output$Type=="Haploid",]
Haploids30C=Haploids[Haploids$Temperature==30,]
Haploids34C=Haploids[Haploids$Temperature==34,]
Haploids37C=Haploids[Haploids$Temperature==37,]

Haploidcontrols=Haploids[Haploids$ERG11_loc=="Native",]
Haploidcontrols30C=Haploidcontrols[Haploidcontrols$Temperature==30,]
Haploidcontrols34C=Haploidcontrols[Haploidcontrols$Temperature==34,]
Haploidcontrols37C=Haploidcontrols[Haploidcontrols$Temperature==37,]
  
Diploids=All_growthcurve_output[All_growthcurve_output$Type=="Diploid",]
Diploid30C=Diploids[Diploids$Temperature==30,]
Diploid34C=Diploids[Diploids$Temperature==34,]
Diploid37C=Diploids[Diploids$Temperature==37,]

HvD=All_growthcurve_output[All_growthcurve_output$ERG11_loc=="Native",]
#HvD=HvD[HvD$Strain!="DGY3627",]
Hvd30C=HvD[HvD$Temperature==30,]
Hvd34C=HvD[HvD$Temperature==34,]
Hvd37C=HvD[HvD$Temperature==37,]

#Look for outliers within each type
ggplot(Haploidcontrols30C,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("G.R. - hap, 30C") + theme_bw()
#J9 is a visible outlier at 16 and 32 ng/uL.

ggplot(Haploidcontrols34C,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("G.R. - hap, 34C") + theme_bw()

ggplot(Haploidcontrols34C,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("AUC - hap, 34C") + theme_bw()
#AUC more realistic at higher temp and FLC conc.
#N21 is a total outlier batch. D5 doesn't look great either. Remove both.

ggplot(Haploidcontrols30C,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("AUC - hap, 30C") + theme_bw()

ggplot(Haploidcontrols37C,aes(x=as.factor(FLC_conc),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("AUC - hap, 37C") + theme_bw()
#All D batches are wildly different than other batches. Remove them all.

ggplot(Haploids30C,aes(x=ERG11_loc,y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~FLC_conc) + ggtitle("AUC - all hap, 30C") + theme_bw()
#Roughly makes sense

ggplot(Haploids34C,aes(x=ERG11_loc,y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~FLC_conc) + ggtitle("AUC - all hap, 34C") + theme_bw()

ggplot(Haploids37C,aes(x=ERG11_loc,y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~FLC_conc) + ggtitle("AUC - all hap, 37C") + theme_bw()
#And again

#Note - the removed high temp batches: N21, and D batches, were Nov-Dec 2025, all made from a single batch of media, which might be why they were all off and different than all other batches.
```
Remove identified outlier batches
```{r}

Filtered_growthcurve_output=All_growthcurve_output[All_growthcurve_output$run!="J9",]
#Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="A26",]
Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="N21",]
Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="D5",]
Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="D10",]
Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="D12",]
Filtered_growthcurve_output=Filtered_growthcurve_output[Filtered_growthcurve_output$run!="D18",]

#separate by type
Haploids=Filtered_growthcurve_output[Filtered_growthcurve_output$Type=="Haploid",]
Haploids30C=Haploids[Haploids$Temperature==30,]
Haploids34C=Haploids[Haploids$Temperature==34,]
Haploids37C=Haploids[Haploids$Temperature==37,]

Haploidcontrols=Haploids[Haploids$ERG11_loc=="Native",]
Haploidcontrols30C=Haploidcontrols[Haploidcontrols$Temperature==30,]
Haploidcontrols34C=Haploidcontrols[Haploidcontrols$Temperature==34,]
Haploidcontrols37C=Haploidcontrols[Haploidcontrols$Temperature==37,]
  
Diploids=Filtered_growthcurve_output[Filtered_growthcurve_output$Type=="Diploid",]
Diploid30C=Diploids[Diploids$Temperature==30,]
Diploid34C=Diploids[Diploids$Temperature==34,]
Diploid37C=Diploids[Diploids$Temperature==37,]

HvD=Filtered_growthcurve_output[Filtered_growthcurve_output$ERG11_loc=="Native",]
Hvd30C=HvD[HvD$Temperature==30,]
Hvd34C=HvD[HvD$Temperature==34,]
Hvd37C=HvD[HvD$Temperature==37,]
```

```{r tempeffect}
ggplot(Haploidcontrols,aes(x=as.factor(Temperature),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~FLC_conc) +ggtitle("AUC - haploid") + theme_bw()

#Haploidcontrols_relconcs=Haploidcontrols[Haploidcontrols$FLC_conc!=4,]
#Haploidcontrols_relconcs=Haploidcontrols_relconcs[Haploidcontrols_relconcs$FLC_conc!=128,]
#ggplot(Haploidcontrols_relconcs,aes(x=as.factor(Temperature),y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~FLC_conc) +ggtitle("AUC - hap - common concs") + theme_bw()
```

```{r heatmap}
HaploidcontrolAUCsummary=aggregate(Haploidcontrols$auc_l,list(Haploidcontrols$FLC_conc, Haploidcontrols$Temperature),mean)
colnames(HaploidcontrolAUCsummary)=c("FLC_conc","Temperature","mean_AUC")

#Make everything relative to 30C/noFLC
HaploidcontrolAUCsummary$rel_AUC=HaploidcontrolAUCsummary$mean_AUC/HaploidcontrolAUCsummary$mean_AUC[1]
#Express it as a percentage
HaploidcontrolAUCsummary$rel_AUC=HaploidcontrolAUCsummary$rel_AUC*100

Hmap=ggplot(HaploidcontrolAUCsummary, aes(x=as.factor(FLC_conc), y=as.factor(Temperature), fill=rel_AUC)) + geom_tile() + scale_fill_gradient(low="khaki1", high="limegreen") + theme_bw() + theme(panel.border=element_blank(), panel.grid.major=element_blank(), panel.grid.minor=element_blank(), axis.text=element_text(face="bold", size=10), axis.title=element_text(face="bold", size=12)) + xlab("FLC conc ng/uL") + ylab("Temperature C")

Hmap
ggsave("./HaploidcontrolAUChmap.pdf", plot = Hmap, device = "pdf")
```



```{r stderror}
#95% CIs are 1.96 X standard deviation / sqrt (# observations). Calculated here from across replicates for each strain/conc, and added as a vertical line to each strain/conc.

Haploidcontrols %>%
  group_by(Temperature,FLC_conc) %>%
  summarise(
    sem=1.96*sd(auc_l)/sqrt(n()),
    auc=mean(auc_l),
  ) -> Meanerror

Meanerror

write.table(Meanerror, "AUC_meanserrors.csv", row.names=FALSE)

ggplot(Meanerror, aes(x=as.factor(FLC_conc), y=auc)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=as.factor(Temperature), color=as.factor(Temperature))) + geom_errorbar(width=0, aes(ymin=auc-sem, ymax=auc+sem)) + ggtitle("Control strain AUC")  + theme_bw()

ggplot(Meanerror, aes(x=FLC_conc, y=auc, color=as.factor(Temperature))) + geom_line() + geom_point() + geom_errorbar(width=0, aes(ymin=auc-sem, ymax=auc+sem)) + ggtitle("Control strain AUC")  + theme_bw()
```

```{r }
#95% CIs are 1.96 X standard deviation / sqrt (# observations). Calculated here from across replicates for each strain/conc, and added as a vertical line to each strain/conc.

HvD %>%
  group_by(Temperature,FLC_conc, Type) %>%
  summarise(
    sem=1.96*sd(auc_l)/sqrt(n()),
    auc=mean(auc_l),
  ) -> Meanerror_HvD

Meanerror_HvD

write.table(Meanerror_HvD, "HvD_AUCmeanserror.csv", row.names=FALSE, sep = "\t")

ggplot(Meanerror_HvD, aes(x=as.factor(FLC_conc), y=auc, color=as.factor(Temperature), shape=Type, linetype=Type)) + geom_point(position=position_dodge(width=0.5), size=3) + geom_errorbar(aes(ymin=auc-sem, ymax=auc+sem), position=position_dodge(width=0.5), width=0.1) + scale_shape_manual(values=c("Haploid"=16, "Diploid"=1)) + scale_color_manual(values=c("30"="yellow3","34"="darkorange","37"="red"), name="Temp. (°C)") + scale_linetype_manual(values=c("Haploid"="solid", "Diploid"="dashed"), name="Background ploidy") + labs(x="FLC conc ng/uL", y="Mean AUC", title="Growth by temp. and type, across concs") + theme_bw()


ggplot(Meanerror_HvD, aes(x=FLC_conc, y=auc, color=as.factor(Temperature), linetype=Type, group=interaction(Temperature, Type))) + geom_line() + geom_point(size=1.5) + geom_errorbar(aes(ymin=auc-sem, ymax=auc+sem), width=0.1) + scale_color_manual(values=c("30"="yellow3","34"="darkorange","37"="red"), name="Temperature (°C)") + scale_linetype_manual(values=c("Haploid"="solid", "Diploid"="dashed"), name="Type") + labs(x="FLC conc ng/uL", y="Mean AUC", title="Growth across FLC, temps, ploidy") + theme_bw()

```