---
title: "Analysis of growth curve outputs"
author: "Saaz Sakrikar"
date: "2026-03-27"
output: html_document
---

```{r setup, warning=FALSE}
library(ggplot2)
library(readxl)
library(forcats)
library(tidyverse)

All_growthcurve_output = read_excel("All_growthcurve_output_2025Apr7.xlsx")
```

Overview plot- all data, all FLC concs, all strains. Then repeat this by removing the FLC concs at extremes (FLC 4 and FLC 128). Finally repeat this by removing the June 9 which had strange control values. Finally show this but separated only by control vs CNV.
```{r overview_plot}
ggplot(All_growthcurve_output,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=Type, color=Type)) + facet_wrap(~Strain) + ggtitle("Growth rate for all strains") + theme_bw()

#Sampledata
Sample_output=All_growthcurve_output[All_growthcurve_output$Strain=="DGY2349"|All_growthcurve_output$Strain=="DGY3268",]

ggplot(Sample_output,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=Type, color=Type)) + facet_wrap(~Strain) + ggtitle("Growth rate for selected strains") + theme_bw()

#Removing FLC4 and FLC 128.
Commonconcs_growth=All_growthcurve_output[All_growthcurve_output$FLC_conc!=4,]
Commonconcs_growth=Commonconcs_growth[Commonconcs_growth$FLC_conc!=128,]

ggplot(Commonconcs_growth,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=Type, color=Type)) + facet_wrap(~Strain) + ggtitle("Growth rate by strain - common FLC concs") + theme_bw()

#Look for outlier batches in the control strains
Control_commonconcs=Commonconcs_growth[Commonconcs_growth$Type=="Control",]
ggplot(Control_commonconcs,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + facet_wrap(~Strain) + ggtitle("Growth rate of control strains - common FLC concs") + theme_bw()

#Removing June 9 batch which is a visible outlier, and plot all remaining data.
Commonconcs_goodbatches=Commonconcs_growth[Commonconcs_growth$run!="J9",]

#ggplot(Commonconcs_goodbatches,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=Type, color=Type)) + facet_wrap(~Strain) + ggtitle("Growth rate by strain - common FLC concs, excl outlier batch") + theme_bw()

#Inspecting control strains after removing outlier batch
Commonconcs_goodbatches_control=Commonconcs_goodbatches[Commonconcs_goodbatches$Type=="Control",]

ggplot(Commonconcs_goodbatches_control,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=run, color=run)) + ggtitle("Growth rate of control strains") + theme_bw()


```

The last graph shows no particular batch effect in the control strains (DGY1315, 500, 3183). This allows us to do comparisons across batches. What I want to do here is to find, at every FLC conc, the ratio of the growth rate of a given CNV strain to that of the avg across all control strains.

```{r ratioplot_prep}

#Avg of control strains across batches
Controlmeans=aggregate(Commonconcs_goodbatches_control$r,list(Commonconcs_goodbatches_control$FLC_conc),mean)

#Getting CNV data separately
Commonconcs_goodbatches_CNV=Commonconcs_goodbatches[Commonconcs_goodbatches$Type!="Control",]
Commonconcs_goodbatches_CNV$Growthrate_ratio=0

#CNVstrainnames=unique(Commonconcs_goodbatches_CNVstrains_output$Strain)

l=length(Commonconcs_goodbatches_CNV$FLC_conc)

for (i in (1:l)) {
  relconc=Commonconcs_goodbatches_CNV$FLC_conc[i]
  relvalue=Controlmeans$x[Controlmeans$Group.1==relconc]
  Commonconcs_goodbatches_CNV$Growthrate_ratio[i]=Commonconcs_goodbatches_CNV$r[i]/relvalue
}

```

Plot the ratio data generated above
```{r ratioplots}
#Plot of each CNV
#ggplot(Commonconcs_goodbatches_CNV,aes(x=as.factor(FLC_conc),y=Growthrate_ratio)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=Strain, color=Strain)) + facet_wrap(~Strain) + ggtitle("CNV strain growth rate ratios") + geom_hline(yintercept=1,linetype=2) + theme_bw()

#Plot of each CNV, coloured by the conc in which CNV was evolved
ggplot(Commonconcs_goodbatches_CNV,aes(x=as.factor(FLC_conc),y=Growthrate_ratio)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=as.factor(FLC_evolved), color=as.factor(FLC_evolved))) + facet_wrap(~Strain) + ggtitle("CNV strain growth rate ratios") + geom_hline(yintercept=1,linetype=2) + theme_bw()

#Plot of each CNV, coloured by the temp in which CNV was evolved
#ggplot(Commonconcs_goodbatches_CNV,aes(x=as.factor(FLC_conc),y=Growthrate_ratio)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=as.factor(Temp_evolved), color=as.factor(Temp_evolved))) + facet_wrap(~Strain) + ggtitle("CNV strain growth rate ratios") + geom_hline(yintercept=1,linetype=2) + theme_bw()

```

```{r stderror}
#95% CIs are calculated using a t-distribution..

Commonconcs_goodbatches_CNV %>%
  group_by(Strain,FLC_conc) %>%
  summarise(
    sem=qt(p=0.95,df=(n()-1))*sd(Growthrate_ratio)/sqrt(n()),
    rate=mean(Growthrate_ratio),
    FLC_evolved=mean(FLC_evolved),
    Temp_evolved=mean(Temp_evolved)
  ) -> t_meanerror_CNV

t_meanerror_CNV

ggplot(t_meanerror_CNV, aes(x=as.factor(FLC_conc), y=rate)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=as.factor(FLC_evolved), color=as.factor(FLC_evolved))) + geom_errorbar(width=0, aes(ymin=rate-sem, ymax=rate+sem)) + facet_wrap(~Strain) + ggtitle("CNV strain growth rate ratios") + geom_hline(yintercept=1,linetype=2) + theme_bw()

write.csv(t_meanerror_CNV,"t_meanerror_CNV.csv")
```

```{r heatmap, fig.show="hold", warning=FALSE, message=FALSE}

#Following code is used to index the data frame, to add a vertical line representing the FLC conc at which each strain was evolved.

Commonconcs_goodbatches_CNV_ordered=Commonconcs_goodbatches_CNV[order(Commonconcs_goodbatches_CNV$Strain,decreasing=TRUE),]
Commonconcs_goodbatches_CNV_ordered=Commonconcs_goodbatches_CNV_ordered[order(Commonconcs_goodbatches_CNV_ordered$FLC_evolved),]

Templist=Commonconcs_goodbatches_CNV_ordered$Strain
Indexing=as.numeric(vector(length=length(Templist)))
Indexing=Indexing+1
for (i in 2:length(Templist)) {
  value=Indexing[i] 
  if (Templist[i-1]!=Templist[i]) { 
    Indexing[i:length(Templist)]=value+1 
  } 
}
Commonconcs_goodbatches_CNV_ordered$indexing=Indexing

#Make a plot with strain names attached for my own reference
#ggplot(Commonconcs_goodbatches_CNV_ordered,aes(x=as.factor(FLC_conc),y=fct_inorder(Strain),fill=Growthrate_ratio)) + geom_tile() + scale_fill_gradient2(low="blue", high="red", mid="black",midpoint = 1) + theme_bw() + theme(strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y=element_blank()) + scale_x_discrete(expand=c(0,0.7)) + geom_segment(aes(x = 0.4, xend = 0.4, y = indexing, yend = c(indexing[-1], max(indexing)), color = as.factor(Temp_evolved)), size = 2) + scale_color_manual(values=c("white","yellow3","darkorange","red"))


ggplot(Commonconcs_goodbatches_CNV_ordered,aes(x=as.factor(FLC_conc),y=fct_inorder(Strain),fill=Growthrate_ratio)) + geom_tile() + scale_fill_gradient2(low="cadetblue1", high="violetred1", mid="black",midpoint = 1) + theme_bw() + theme(strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y=element_blank()) + scale_x_discrete(expand=c(0,0.7)) + geom_segment(aes(x = 0.4, xend = 0.4, y = indexing, yend = c(indexing[-1], max(indexing)), color = as.factor(FLC_evolved)), size = 2) + scale_color_manual(values=c("white","bisque2","bisque4","black")) 

#Plot for publication


Heatmap=ggplot(Commonconcs_goodbatches_CNV_ordered,aes(x=as.factor(FLC_conc),y=fct_inorder(Strain),fill=Growthrate_ratio)) + geom_tile() + scale_fill_gradient2(low="navyblue", high="violetred1", mid="white",midpoint = 1) + theme_bw() + theme(strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y=element_blank()) 


Heatmap
#ggsave("./Heatmap_allstrainsgrowth.tiff", plot = Heatmap, device = "tiff")
#ggsave("./Heatmap_allstrainsgrowth.pdf", plot = Heatmap, device = "pdf")
```