---
title: "Graphs showing trajectory of ERG11 CNVs"
author: "Saaz Sakrikar"
date: "2025-6-6"
output: html_document
---

Mostly hard-coded and self-explanatory.

Script "plotprocessing_passage.R" is used here.
```{r libraries,warning=FALSE,message=FALSE}
library(readxl)
library(ggplot2)
library(reshape2)
library(gghighlight)
library(forcats)
library(RColorBrewer)
library(ggh4x)
source("plotprocessing_passage.R")
```

Load each tab of my excel sheet.

```{r load_data}
Jun12_30C=read_excel("Apr2025_allruns.xlsx",sheet="30C_June12")

Dec22_30C=read_excel("Apr2025_allruns.xlsx",sheet="30C_Dec22")

Nov1_34C=read_excel("Apr2025_allruns.xlsx",sheet="34C_Nov1")

Sep20_37C=read_excel("Apr2025_allruns.xlsx",sheet="37C_Sep20")

Dec22_37C=read_excel("Apr2025_allruns.xlsx",sheet="37C_Dec22")
```

Data processing: making plottable dataframes using melt from reshape, adding a column of "genotype" to allow grouping of lines by genotype. All of this is done within a script.

```{r initial_processing}
Jun12_30C_melted = plotprocessing_passage(Jun12_30C)
colnames(Jun12_30C_melted)=c("Passage","variable","value","FLC_conc")

Dec22_30C_melted = plotprocessing_passage(Dec22_30C)
colnames(Dec22_30C_melted)=c("Passage","variable","value","FLC_conc")

Nov1_34C_melted = plotprocessing_passage(Nov1_34C)
colnames(Nov1_34C_melted)=c("Passage","variable","value","FLC_conc")

Sep20_37C_melted = plotprocessing_passage(Sep20_37C)
colnames(Sep20_37C_melted)=c("Passage","variable","value","FLC_conc")

Dec22_37C_melted = plotprocessing_passage(Dec22_37C)
colnames(Dec22_37C_melted)=c("Passage","variable","value","FLC_conc")
```


```{r combine}
PasslistJ30=vector(length=length(Jun12_30C_melted$Passage))
l=length(PasslistJ30)
PasslistJ30[1:l]=30
Jun12_30C_melted$Temp=PasslistJ30
Jun12_30C_melted$Batch="J12"

PasslistD30=vector(length=length(Dec22_30C_melted$Passage))
l=length(PasslistD30)
PasslistD30[1:l]=30
Dec22_30C_melted$Temp=PasslistD30
Dec22_30C_melted$Batch="D22_30"

PasslistN34=vector(length=length(Nov1_34C_melted$Passage))
l=length(PasslistN34)
PasslistN34[1:l]=34
Nov1_34C_melted$Temp=PasslistN34
Nov1_34C_melted$Batch="N1"

PasslistS37=vector(length=length(Sep20_37C_melted$Passage))
l=length(PasslistS37)
PasslistS37[1:l]=37
Sep20_37C_melted$Temp=PasslistS37
Sep20_37C_melted$Batch="S20"

PasslistD37=vector(length=length(Dec22_37C_melted$Passage))
l=length(PasslistD37)
PasslistD37[1:l]=37
Dec22_37C_melted$Temp=PasslistD37
Dec22_37C_melted$Batch="D22_37"

Allruns_bypass=rbind(Jun12_30C_melted, Dec22_30C_melted, Nov1_34C_melted, Sep20_37C_melted, Dec22_37C_melted)
Allruns_bypass$sample=paste(Allruns_bypass$variable,"_",Allruns_bypass$Temp)
Allruns_bypass$Condition=paste(Allruns_bypass$Temp,"_",Allruns_bypass$FLC_conc)
Allruns_bypass$Temp=as.factor(Allruns_bypass$Temp)
```

```{r combined_graphs,fig.show="hold",message=FALSE,warning=FALSE}

facet_design= c(
    "
    AABBCC
    DDEEFF
    ##GG##
    "
)
Bypassage_smoothed_figure = ggplot(Allruns_bypass,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) +  facet_manual(~fct_inseq(FLC_conc),design=facet_design) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("Emergence of ERG11 CNVs") + theme(text=element_text(size=12,face="bold"),) + ylab("CNV %") + scale_color_manual(values=c("yellow3","darkorange","red")) + xlab("# passages")

Bypassage_smoothed_figure
#ggsave("./Bypassage_smoothed_figure.tiff", plot = Bypassage_smoothed_figure, device = "tiff")
#ggsave("./Bypassage_smoothed_figure.pdf", plot = Bypassage_smoothed_figure, device = "pdf")

#Splitting by both temperature and FLC conc
facet_design_2split=c("
                AABBCC
                DD##EE
                FF##GG
                HHIIJJ
                KKLLMM
                NNOOPP
                QQRR##")

Allruns_bypass=Allruns_bypass[order(as.numeric(Allruns_bypass$FLC_conc)),]

Smoothed_figure_2split = ggplot(Allruns_bypass,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) +  facet_manual(~fct_inorder(Condition),design=facet_design_2split) + geom_smooth(se=FALSE) + theme_bw() + theme(text=element_text(size=11,face="bold"), strip.background=element_blank(), strip.text.x=element_blank()) + ylab("CNV %") + scale_color_manual(values=c("yellow3","darkorange","red")) + xlab("# passages")

Smoothed_figure_2split
#ggsave("./Smoothed_figure_2split.tiff", plot = Smoothed_figure_2split, device = "tiff")
#ggsave("./Smoothed_figure_2split.pdf", plot = Smoothed_figure_2split, device = "pdf")
```

Show temperature dependence by making graphs across temps at FLC 32.
```{r ppt_tempgraphs, fig.show="hold",warning=FALSE,message=FALSE}

FLC32_bypass_allruns=Allruns_bypass[Allruns_bypass$FLC_conc=="32",]

FLC32_bypass_figure=ggplot(FLC32_bypass_allruns,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~FLC_conc) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("Strong temperature dependence of CNV dynamics") + theme(text=element_text(size=12,face="bold"),) + ylab("Two copy %") + scale_color_manual(values=c("yellow3","darkorange","red")) + ylim(c(0,100)) + xlab("#passages")
FLC32_bypass_figure
#ggsave("./FLC32_bypass.tiff", plot = FLC32_bypass_figure, device = "tiff")
#ggsave("./FLC32_bypass.pdf", plot = FLC32_bypass_figure, device = "pdf")

FLC32_bypass_30Cruns=FLC32_bypass_allruns[FLC32_bypass_allruns$Temp=="30",]

FLC32_bypass_30Conly=ggplot(FLC32_bypass_30Cruns,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~FLC_conc) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("No emergence of ERG11 CNVs in 32ng/uL FLC at 30C") + theme(text=element_text(size=12,face="bold"),) + ylab("Two copy %") + scale_color_manual(values=c("yellow3","darkorange","red")) + ylim(c(0,100)) + xlab("# passages")
FLC32_bypass_30Conly
#ggsave("./FLC32_bypass_30Conly.tiff", plot = FLC32_bypass_30Conly, device = "tiff")
#ggsave("./FLC32_bypass_30Conly.pdf", plot = FLC32_bypass_30Conly, device = "pdf")

FLC32_bypass_no37runs=FLC32_bypass_allruns[FLC32_bypass_allruns$Temp!="37",]

FLC32_bypass_no37=ggplot(FLC32_bypass_no37runs,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~FLC_conc) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("Strong selection for CNVs after small temp. increase") + theme(text=element_text(size=12,face="bold"),) + ylab("Two copy %") + scale_color_manual(values=c("yellow3","darkorange","red")) + ylim(c(0,100)) + xlab("# passages")
FLC32_bypass_no37
#ggsave("./FLC32_bypass_no37C.tiff", plot = FLC32_bypass_no37, device = "tiff")
#ggsave("./FLC32_bypass_no37C.pdf", plot = FLC32_bypass_no37, device = "pdf")
```

Show concentration dependence - across 3 concs at 30ng/uL

```{r ppt_concgraphs}
C30_bypass_allruns=Allruns_bypass[Allruns_bypass$Temp=="30",]

C30_bypass_relruns=C30_bypass_allruns[C30_bypass_allruns$FLC_conc=="8"| C30_bypass_allruns$FLC_conc=="16" | C30_bypass_allruns$FLC_conc=="32",]

C30_bypass_figure=ggplot(C30_bypass_relruns,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~fct_inorder(FLC_conc)) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("Emergence of ERG11 CNVs in 30C at lower concs") + theme(text=element_text(size=12,face="bold"),) + ylab("Two copy %") + scale_color_manual(values=c("yellow3","darkorange","red")) + ylim(c(0,100)) + xlab("#passages")
C30_bypass_figure
#ggsave("./30C_bypass_figure.tiff", plot = C30_bypass_figure, device = "tiff")
#ggsave("./30C_bypass_figure.pdf", plot = C30_bypass_figure, device = "pdf")

# C30_bypass_figure=FLC32_bypass_allruns[FLC32_bypass_allruns$Temp=="30",]
# 
# FLC32_bypass_30Conly=ggplot(FLC32_bypass_30Cruns,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~FLC_conc) + geom_smooth(se=FALSE) + theme_bw() + ggtitle("No emergence of ERG11 CNVs in 32ng/uL FLC at 30C") + theme(text=element_text(size=12,face="bold"),) + ylab("Two copy %") + scale_color_manual(values=c("yellow3","darkorange","red")) + ylim(c(0,100)) + xlab("# passages)")
# FLC32_bypass_30Conly
#ggsave("./FLC32_bypass_30Conly.tiff", plot = FLC32_bypass_30Conly, device = "tiff")
#ggsave("./FLC32_bypass_30Conly.pdf", plot = FLC32_bypass_30Conly, device = "pdf")


```

Remake this figure in all the specific FLC concs (8, 16, 32) where CNVs were actually observed.


```{r Selected_concs, fig.height=7, fig.width=12}

Allruns_bypass_selectedconc=Allruns_bypass
Allruns_bypass_selectedconc$FLC_conc=as.numeric(Allruns_bypass_selectedconc$FLC_conc)
Allruns_bypass_selectedconc=Allruns_bypass_selectedconc[Allruns_bypass_selectedconc$FLC_conc>4,]
Allruns_bypass_selectedconc=Allruns_bypass_selectedconc[Allruns_bypass_selectedconc$FLC_conc<64,]
Allruns_bypass_selectedconc$FLC_conc=as.factor(Allruns_bypass_selectedconc$FLC_conc)

Evolved_bypass_smoothed = ggplot(Allruns_bypass_selectedconc,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~fct_rev(fct_inseq(FLC_conc)),ncol=1) + geom_smooth(se=FALSE) + theme_bw() + theme(text=element_text(size=12,face="bold"),) + ylab("CNV %") + scale_color_manual(values=c("yellow3","darkorange","red")) + coord_fixed(ratio=0.15) + xlab("# passages")

Evolved_bypass_smoothed
#ggsave("./Evolved_bypass.tiff", plot = Evolved_bypass_smoothed, device = "tiff")
ggsave("./Evolved_bypass.pdf", plot = Evolved_bypass_smoothed, device = "pdf")

Evolved_bypass_truncated = ggplot(Allruns_bypass_selectedconc,aes(x=Passage,y=value,group=fct_inorder(sample),color=Temp)) + facet_wrap(~fct_rev(fct_inseq(FLC_conc)),ncol=1) + geom_smooth(se=FALSE) + theme_bw() + scale_x_continuous(limits=c(0,8)) + theme(text=element_text(size=12,face="bold"),) + ylab("CNV %") + scale_color_manual(values=c("yellow3","darkorange","red")) + coord_fixed(ratio=0.18) + xlab("# passages")

Evolved_bypass_truncated
#ggsave("./Evolved_bypass_truncated.tiff", plot = Evolved_bypass_truncated, device = "tiff")
#ggsave("./Evolved_bypass_truncated.pdf", plot = Evolved_bypass_truncated, device = "pdf")

```

Display evolution data as a heatmap rather than a line plot. Quick visual inspection possible.
```{r heatmap, fig.show="hold",warning=FALSE,message=FALSE}

Allruns_bypass_ordered=Allruns_bypass[order(as.numeric(Allruns_bypass$FLC_conc),as.numeric(Allruns_bypass$Temp)),]

#Code that creates an index of samples, which is used to add a vertical line to the heatmap. This vertical line is colour-coded based on this #indexing and the temperature of evolution of the samples, giving  a #guide to the data without needing extra words. 

Tempsamples=Allruns_bypass_ordered$sample
Indexing=as.numeric(vector(length=length(Tempsamples)))
Indexing=Indexing+1

for (i in 2:length(Tempsamples)) {
  value=Indexing[i]
  if (Tempsamples[i-1]!=Tempsamples[i]) {
    Indexing[i:length(Indexing)]=value+1 
  }
}

Allruns_bypass_ordered$indexing=Indexing

Heatmap = ggplot(Allruns_bypass_ordered,aes(x=Passage,y=fct_inorder(sample),fill=value)) + geom_tile(color="white", lwd=0.5) + scale_fill_gradient(low="aquamarine", high="navyblue", na.value="white") + theme_bw() + scale_x_continuous(expand=c(0.005,0)) + theme(strip.background = element_blank(),strip.text.x = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) + geom_segment(aes(x = -0.5, xend = -0.5, y = indexing, yend = c(indexing[-1], max(indexing)), color = Temp), size = 2) + geom_segment(aes(x = -0.75, xend = -0.75, y = indexing, yend = c(indexing[-1], max(indexing)), color = as.factor(FLC_conc)), size = 2) + scale_color_manual(values=c("yellow3","darkorange","red","beige","black", "bisque2","bisque3","blanchedalmond","bisque4","bisque1")) 

Heatmap
#ggsave("./Heatmap_evolutions.tiff", plot = Heatmap, device = "tiff")
#ggsave("./Heatmap_evolutions.pdf", plot = Heatmap, device = "pdf")

Heatmap_truncated = ggplot(Allruns_bypass_ordered,aes(x=Passage,y=fct_inorder(sample),fill=value)) + geom_tile(color="white", lwd=0.45) + scale_fill_gradient(low="aquamarine", high="navyblue", na.value="grey80") + theme_bw()  + scale_x_continuous(limits=c(-0.7,8.5), expand=c(0,0.05)) + theme(strip.background = element_blank(),strip.text.x = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) + geom_segment(aes(x = -0.6, xend = -0.6, y = indexing, yend = c(indexing[-1], max(indexing)), color = Temp), size = 3) + scale_color_manual(values=c("yellow3","darkorange","red")) + coord_fixed(ratio=0.12)

Heatmap_truncated
#ggsave("./Heatmap_8pass.tiff", plot = Heatmap_truncated, device = "tiff")
#ggsave("./Heatmap_8pass.pdf", plot = Heatmap_truncated, device = "pdf")

```

```{r batches}
C30_bypass_relruns2=C30_bypass_relruns[C30_bypass_relruns$FLC_conc=="16",]


ggplot(C30_bypass_relruns2,aes(x=Passage,y=value,group=fct_inorder(variable),color=Batch)) + facet_wrap(~fct_inseq(FLC_conc)) +  geom_smooth(se=FALSE) + theme_bw() + geom_point() + ggtitle("ERG11 CNVs in haploid S cer 30C") + theme(text=element_text(size=12,face="bold"), legend.position="none") + ylab("Two copy %")

C37_bypass_allruns=Allruns_bypass[Allruns_bypass$Temp=="37",]
C37_bypass_relruns=C37_bypass_allruns[C37_bypass_allruns$FLC_conc=="16" | C37_bypass_allruns$FLC_conc=="32",]

ggplot(C37_bypass_relruns,aes(x=Passage,y=value,group=fct_inorder(variable),color=Batch)) + facet_wrap(~fct_inseq(FLC_conc)) +  geom_smooth(se=FALSE) + theme_bw() + geom_point() + ggtitle("ERG11 CNVs in haploid S cer 37C") + theme(text=element_text(size=12,face="bold"), legend.position="none") + ylab("Two copy %")
```
