---
title: "Depth_visualisation_comparison_March17"
author: "Saaz Sakrikar"
date: "2025-03-17"
output: html_document
---

A way of visualising read depth across all chromosomes in yeast, adapted from the method by Biostars user "cmdcolin" in https://www.biostars.org/p/475162.

My depth data is from bedtools genomecov and includes a reading at each individual base, and is quite different from the bed file generated by mosdepth in the above example.

This code visually compares the genome-wide depth from an ancestral strain, to the evolved strain continaing a CNV.

```{r setup}
library(tidyverse)
library(zoo)

chr_sizes=read.table("GCA_000146045.2_R64_Chr8ERG11_July20.fa.fai")
depth_ancestral=read.table("3183_July20_perbase.txt")
depth_CNV=read.table("3227_Oct16_perbase.txt")
```


```{r location_processing}

chr_sizes=chr_sizes[,c(1,2)]
colnames(chr_sizes)=c("Chr","Size")
colnames(depth_ancestral)=c("Chr","Loc","Depth")
colnames(depth_CNV)=c("Chr","Loc","Depth")

#Get a cumulative locus coordinate, which would run along the whole genome rather than each individual chromosome.
cumsize=cumsum(chr_sizes$Size)
cumsize=cumsize[1:length(cumsize)-1]
cumsize=c(0,cumsize)
chr_sizes$cumsize=cumsize

comb_df_ancestral=inner_join(depth_ancestral,chr_sizes,by="Chr")
comb_df_ancestral$cum_loc=comb_df_ancestral$Loc+comb_df_ancestral$cumsize

comb_df_CNV=inner_join(depth_CNV,chr_sizes,by="Chr")
comb_df_CNV$cum_loc=comb_df_CNV$Loc+comb_df_CNV$cumsize
```

Remove rDNA locus on Chr XII, which is known to have variable copy number.

```{r rDNA}
#rDNA exclusion
pos1=451450
pos2=468950

rDNAchr=chr_sizes$Chr[12]
pos1=which(comb_df_ancestral$Chr==rDNAchr & comb_df_ancestral$Loc==pos1)
pos2=which(comb_df_ancestral$Chr==rDNAchr & comb_df_ancestral$Loc==pos2)

comb_df_ancestral=comb_df_ancestral[-c(pos1:pos2),]
comb_df_CNV=comb_df_CNV[-c(pos1:pos2),]
  
```

By trial-and-error, I found that there is a variable region on Chr VIII, between 215000 and 220000 on that chromosome. It contains several genes, most interestingly CUP1-2, CUP1-1, and RSC30, as well as MED6 and CPR2. It seems likely that the variability in this region is due to the presence of CUP1. I will therefore exclude this region from the visualisation, just like with the rDNA locus on Chr 12.

```{r Chr8var}
pos3=215000
pos4=220000

varchr=chr_sizes$Chr[8]
pos3=which(comb_df_ancestral$Chr==varchr & comb_df_ancestral$Loc==pos3)
pos4=which(comb_df_ancestral$Chr==varchr & comb_df_ancestral$Loc==pos4)

comb_df_ancestral=comb_df_ancestral[-c(pos3:pos4),]
comb_df_CNV=comb_df_CNV[-c(pos3:pos4),]

```

Processing the data for visualisation:
Bin the data into 1kb chunks (within each chr) and get avg depth of each chunk. Then taking a rolling mean of the resulting data.
The 1kb chunks reduces the size of the data by 3 order of magnitude, making visualisation possible. The rolling mean removes insane outliers.
```{r depth_processing,warning=FALSE}

#Normalise depth
reldepth_ancestral=comb_df_ancestral$Depth/mean(comb_df_ancestral$Depth)
reldepth_CNV=comb_df_CNV$Depth/mean(comb_df_CNV$Depth)

comb_df_ancestral$reldepth=reldepth_ancestral
comb_df_CNV$reldepth=reldepth_CNV

#Bin data into 1kb chunks; get the avg depth of each chunk. This NEEDS to be done chr-by-chr to avoid messing up coordinates at either end of "adjacent" chromosomes.

Chrnames=unique(comb_df_ancestral$Chr)
l=length(Chrnames)

Newcomb_df_ancestral=comb_df_ancestral
Newcomb_df_ancestral=Newcomb_df_ancestral[-c(1:length(Newcomb_df_ancestral$Chr)),]

for (i in (1:l)) {
  relchr=Chrnames[i]
  relchrdata=comb_df_ancestral[comb_df_ancestral$Chr==relchr,]
  relchrdata=aggregate(relchrdata,by=list(cut(relchrdata$Loc,seq(0,max(relchrdata$Loc),100))),mean)
  relchrdata=relchrdata[,-1]
  relchrdata$Chr=relchr
  Newcomb_df_ancestral=rbind(Newcomb_df_ancestral,relchrdata)
}

#Get rolling average
rollmeandepth_ancestral=rollmean(Newcomb_df_ancestral$reldepth,1000,fill=1)
Newcomb_df_ancestral$rollmeandepth=rollmeandepth_ancestral


Newcomb_df_CNV=comb_df_CNV
Newcomb_df_CNV=Newcomb_df_CNV[-c(1:length(Newcomb_df_CNV$Chr)),]

for (i in (1:l)) {
  relchr=Chrnames[i]
  relchrdata=comb_df_CNV[comb_df_CNV$Chr==relchr,]
  relchrdata=aggregate(relchrdata,by=list(cut(relchrdata$Loc,seq(0,max(relchrdata$Loc),100))),mean)
  relchrdata=relchrdata[,-1]
  relchrdata$Chr=relchr
  Newcomb_df_CNV=rbind(Newcomb_df_CNV,relchrdata)
}

#Get rolling average
rollmeandepth_CNV=rollmean(Newcomb_df_CNV$reldepth,1000,fill=1)
Newcomb_df_CNV$rollmeandepth=rollmeandepth_CNV

```

Renaming chromosomes - this is trivially easy in this case because the inputs have all chromosomes in order
```{r chrnames}
Oldchrnames=unique(chr_sizes$Chr)

Newchrnames=c("Chr I", "Chr II", "Chr III", "Chr IV", "Chr V", "Chr VI", "Chr VII", "Chr VIII", "Chr IX", "Chr X", "Chr XI", "Chr XII", "Chr XIII", "Chr XIV", "Chr XV", "Chr XVI")

Allchrnames=Newcomb_df_ancestral$Chr
l1=length(Allchrnames)

for (i in (1:l1)) {
  chrnamecurrent=Allchrnames[i]
  relindex=which(Oldchrnames==chrnamecurrent)
  chrnamenew=Newchrnames[relindex]
  Allchrnames[i]=chrnamenew
}

Newcomb_df_ancestral$Newchr=Allchrnames
Newcomb_df_CNV$Newchr=Allchrnames
```

Individual plots
```{r plot, fig.width=7, fig.height=2.5}

ggplot(data=Newcomb_df_ancestral) + geom_line(aes(x=cum_loc, y=rollmeandepth), color="darkgrey") + ylim(c(0.5,2.1)) + scale_x_continuous(label = Newcomb_df_ancestral$Newchr, breaks = Newcomb_df_ancestral$cumsize) + theme_bw() + theme(axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5),axis.title.x=element_blank()) + ylab ("Normalised read depth")

ggplot(data=Newcomb_df_CNV) + geom_line(aes(x=cum_loc, y=rollmeandepth), color="red") + ylim(c(0.5,2.1)) + scale_x_continuous(label = Newcomb_df_ancestral$Newchr, breaks = Newcomb_df_ancestral$cumsize) + theme_bw() + theme(axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5),axis.title.x=element_blank()) + ylab ("Normalised read depth")

```


Combined graphs
```{r combined_plot,fig.show="hold",message=FALSE,warning=FALSE}
Bothstrains_df=as.data.frame(matrix(nrow=nrow(Newcomb_df_ancestral),ncol=5))
colnames(Bothstrains_df)=c("cum_loc","cumsize","Newchr","rmd_ancestral","rmd_CNV")

Bothstrains_df$cum_loc=Newcomb_df_ancestral$cum_loc
Bothstrains_df$cumsize=Newcomb_df_ancestral$cumsize
Bothstrains_df$Newchr=Newcomb_df_ancestral$Newchr
Bothstrains_df$rmd_ancestral=Newcomb_df_ancestral$rollmeandepth
Bothstrains_df$rmd_CNV=Newcomb_df_CNV$rollmeandepth

#Readdepth_straincomparison_full = ggplot(data=Bothstrins_df) + geom_line(aes(x=cum_loc, y=rmd_ancestral), color="darkgrey") + geom_line(aes(x=cum_loc, y=rmd_CNV), color="black") + ylim(c(0,2.2)) + scale_x_continuous(label = Bothstrains_df$Newchr, breaks = Bothstrains_df$cumsize) + theme_bw() + theme(axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5),axis.title.x=element_blank()) + ylab ("Normalised read depth")

#ggsave("./Readdepth_straincomparison_full.tiff", plot = Readdepth_straincomparison_full, device = "tiff")
#ggsave("./Readdepth_straincomparison_full.pdf", plot = Readdepth_straincomparison_full, device = "pdf")

Readdepth_straincomparison = ggplot(data=Bothstrains_df) + geom_line(aes(x=cum_loc, y=rmd_ancestral), color="darkgrey") + geom_line(aes(x=cum_loc, y=rmd_CNV), color="black") + scale_x_continuous(label = Bothstrains_df$Newchr, breaks = Bothstrains_df$cumsize) + theme_bw() + theme(axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5),axis.title.x=element_blank()) + ylab ("Normalised read depth")

#ggsave("./Readdepth_straincomparison.tiff", plot = Readdepth_straincomparison, device = "tiff")
ggsave("./Readdepth_straincomparison.pdf", plot = Readdepth_straincomparison, device = "pdf")

Readdepth_straincomparison
#Readdepth_straincomparison_full
```